On existence of integral point sets and their diameter bounds \footnote{ This work was carried out at Voronezh State University and supported by the Russian Science Foundation grant 19-11-00197. }

### 2020-01-02 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Proof of Lemma 3.4 (ex Lemma 2.4)
As requested in Review 1, p. 1 and Review 2, medium remark 4.

>>>>

\begin{proof}
	Consider $M_1, M_2 \in M$ such that $|M_1 - M_2| = d$
	and a Cartesian coordinate system such that
	$M_1 = (-d/2; 0)$, $M_2 = (d/2; 0)$.

	Then for $M_i =(x_i, y_i)\in M$, $i > 2$,
	the inequality $|x_i| < d/2$ holds
	(otherwise we get $|M_i - M_1| > d$ or $|M_i - M_2| > d$).
	Moreover, $|y_i| < d$ (otherwise we get $|M_i - M_1| > d$ or $|M_i - M_2| > d$).

	Set $y_{\max} = \max_{i} y_i$, $y_{\min} = \min_{i} y_i$, $M_{\max}=(x_{\max}, y_{\max})$, $M_{\min}=(x_{\min}, y_{\min})$
	(in case of multiple points providing the maximum or the minimum, the choice is arbitrary).
	Then
	\begin{multline}
		d \geq |M_{\max} - M_{\min}| = \sqrt{(x_{\max} - x_{\min})^2 + (y_{\max} - y_{\min})^2}
		\geq
		\\\geq
		\sqrt{(y_{\max} - y_{\min})^2} =
		y_{\max} - y_{\min}
	\end{multline}
	Thus, the sides of the square that contains $M$ are $x=\pm d/2$,
	$y=y_{\max}$ and $y=y_{\max} - d$.
\end{proof}
